\section{Methods}\label{sec6}

\subsection{Study design and analytic scope}


This study analyses naturalistic public human--LLM conversations as
observational digital traces of everyday work and study, following the logic of
large-scale digital-trace and computational social-science analyses~\cite{lazer2009computational,salganik2017bitbybit}.
We use three public conversational corpora: WildChat-4.8M~\cite{allenai2025wildchat48m,zhao2024wildchat}
(public non-toxic release, reported below as WildChat), LMSYS Chat-1M~\cite{zheng2023lmsyschat1m} (reported below as LMSYS Chat) and ShareChat~\cite{yan2025sharechat} (strict-English subset, reported below as ShareChat). Within these corpora, we
focus on coding and writing, two prominent AI-mediated knowledge-work ecologies in which users can delegate task completion to the assistant or remain actively involved in diagnosis, evaluation and revision. Usage and workplace
studies identify writing and software-development tasks as high-salience domains of everyday LLM use and AI-mediated knowledge work~\cite{chatterji2025people,handa2025economic,noy2023experimental,brynjolfsson2025generative}. The two domains also differ in how they make learning-oriented participation
visible: coding often externalizes uncertainty through errors, constraints, execution failures and iterative testing~\cite{barke2023grounded,keuning2018systematic},
whereas writing often involves drafting, revision and evaluative choice whose rationale may remain implicit unless users articulate it~\cite{flower1981cognitive,sommers1980revision}.

The design is descriptive and associational. We treat conversations as naturally occurring records of user--assistant collaboration and analyse them at three scales: conversations, turns and adjacent user--assistant--user sequences. User
turns provide engagement-depth measures, assistant turns provide support measures and adjacent sequences provide local ordering. The analyses estimate where learning-oriented engagement is visible, where constructive engagement is
concentrated and how scaffolded support is associated with further constructive participation across conversation-level, support-form and adjacent-turn scales. Dataset and task ecology are treated as substantive dimensions of the interaction
environment. The claims are therefore about observable behavioural signatures and interactional organization in public logs, with causal and durable-learning questions reserved for experimental or longitudinal designs~\cite{hernan2020causal}.


% This study analyses naturalistic public human--LLM conversations as observational digital traces of everyday work and study. We use three public conversational corpora: WildChat-4.8M~\cite{allenai2025wildchat48m} (public non-toxic release, reported below as WildChat), LMSYS Chat-1M~\cite{zheng2023lmsyschat1m} (reported below as LMSYS Chat) and the ShareChat~\cite{yan2025sharechat} (strict-English subset, reported below as ShareChat). Within these corpora, we focus on two common task ecologies: coding and writing. They were selected as prominent AI-mediated knowledge-work settings in which users can either delegate task completion to the assistant or remain actively involved in diagnosis, evaluation and revision. Usage studies identify writing and software-development tasks as high-salience domains of everyday LLM use~\cite{chatterji2025people,handa2025economic,noy2023experimental,brynjolfsson2025generative}. They also differ in how they make learning-oriented participation visible: coding often externalizes uncertainty through errors, constraints, execution failures and iterative testing~\cite{barke2023grounded,keuning2018systematic}, whereas writing often involves drafting, revision and evaluative choice, but the rationale for a change may remain implicit unless users articulate it~\cite{flower1981cognitive,sommers1980revision}.




% The design is descriptive and associational, following the logic of large-scale digital trace and computational social-science analyses~\cite{lazer2009computational,salganik2017bitbybit}. The empirical sequence mirrors the Results section. We first describe whether learning-oriented engagement is visible in ordinary use, then ask where constructive engagement is concentrated, whether scaffolded assistant support marks richer constructive participation, which support forms carry that association and how support and engagement are ordered at the adjacent-turn scale. These analyses support claims about observable behavioural signatures in conversation logs. They do not estimate randomized exposure effects, durable learning outcomes, changes in users' knowledge or stable learner traits.



\subsection{Corpus construction and task settings}




We constructed the analytic corpus from three public human--LLM conversation releases: WildChat-4.8M~\cite{allenai2025wildchat48m,zhao2024wildchat}, LMSYS Chat-1M~\cite{zheng2023lmsyschat1m} and ShareChat~\cite{yan2025sharechat}. For each source, we converted the public records into role-ordered conversation records with user and assistant turns. We then applied corpus-specific language and minimum-turn screens to retain conversations within the English-oriented analysis scope and with at least four message turns. Coding- and writing-oriented conversations were identified with LLM-assisted semantic task filters. Retained conversations were then processed with the shared annotation pipeline for conversation-level user framing and turn-level labels for user engagement, scaffolded support, support intent and support form.



This construction process yielded six analytic task settings: WildChat coding, WildChat writing, LMSYS Chat coding, LMSYS Chat writing, ShareChat coding and ShareChat writing. The final corpus contains 128,569 conversations and 981,470 total turns, including 491,685 user turns and 489,785 assistant turns. WildChat and LMSYS Chat contribute the largest public interaction logs, whereas ShareChat contributes a smaller complementary sample. 

Supplementary Table~\ref{tab:corpus_filtering_pipeline} reports the corpus-specific raw-to-final filtering ledger, including source records, eligibility screens, task-filtering steps and final analytic samples. Supplementary Table~\ref{tab:corpus_provenance_artifacts} records retained provenance artifacts and script-hash prefixes. Supplementary Table~\ref{tab:filter_classifier_validation} summarizes validation and audit evidence for task, language, user-framing and turn-level labels.




% Supplementary Table~\ref{tab:corpus_filtering_pipeline} reports the raw-to-final filtering ledger; Supplementary Table~\ref{tab:corpus_provenance_artifacts} records the retained provenance artifacts and script-hash prefixes; and Supplementary Table~\ref{tab:filter_classifier_validation} summarizes validation and audit evidence for the task, language, user-framing and turn-level labels.

% The analytic corpus contains 128,569 conversations and 981,470 total turns, including 491,685 user turns and 489,785 assistant turns. The six task settings are WildChat coding, WildChat writing, LMSYS Chat coding, LMSYS Chat writing, ShareChat coding and ShareChat writing. WildChat and LMSYS Chat provide large public human--LLM interaction logs from deployed chat systems, whereas ShareChat contributes a smaller strict-English analytic subset after our language screen. The settings are complementary naturalistic environments rather than experimentally matched platform conditions~\cite{zhao2024wildchat,zheng2023lmsyschat1m,yan2025sharechat}.



% Corpus construction followed the same high-level sequence across datasets. Public source records were converted to role-ordered user--assistant conversations; conversations were restricted to English-oriented exchanges with at least four turns; coding and writing task ecologies were assigned with LLM-assisted semantic filters; and retained conversations were passed through the shared annotation pipeline. When keyword prefilters were used, the LLM-assisted semantic task filter retained conversations with a score of at least 0.70. ShareChat additionally underwent a strict-English screen after task filtering, retaining only conversations whose final conversation-level language flag was English. Supplementary Tables~\ref{tab:corpus_provenance_artifacts} and \ref{tab:filter_classifier_validation} provide the raw-to-final provenance ledger, classifier sources and validation evidence for these filtering decisions.


\subsection{Operational constructs}

% \subsubsection{Task ecology, user framing and depth}

% Each conversation is represented as an ordered sequence of user and assistant turns together with available metadata. The main contextual variables are data source, task ecology, user framing, topic and language. Task ecology and user framing are analytic variables generated by the filtering and annotation pipeline; topic, language and model/source metadata are retained where available. User framing is a conversation-level LLM-assisted label for whether the opening user request or early exchange explicitly signals understanding, exploration or learning, as opposed to a request without such explicit learning-oriented framing. We interpret this as an observed textual signal, not as a stable measure of motivation or learner identity.

% Interaction depth is measured from the number and ordering of turns in the conversation. It is used descriptively, as a factor of interactional opportunity and sustained exchange, not as an independent manipulation. Post-answer depth is measured as the number of turns after the first assistant response. Because any-event outcomes are easier to observe in longer conversations, the supplementary analyses include rate-based sensitivity checks for constructive engagement.

\subsubsection{Task ecology, user framing and interaction depth}

Each conversation is represented as an ordered sequence of user and assistant turns together with available metadata. The main contextual variables are data source, task ecology, user framing, topic and language. Task ecology and user framing are analytic variables generated by the corpus-construction and annotation pipeline; topic, language and model/source metadata are retained where available. User framing is a conversation-level LLM-assisted label that captures whether the
opening request or early exchange positions the interaction as an occasion for understanding, exploration or learning, rather than only for completing the immediate task.

Interaction depth summarizes how much exchange unfolded after the initial request. We use turn counts and turn ordering as descriptive measures of opportunity for follow-up, repair and elaboration, rather than as manipulated exposures. Post-answer
depth is measured as the number of turns after the first assistant response. When analyses use whether a conversation contains at least one constructive turn, interaction
depth is interpreted as a visibility condition: longer exchanges provide more occasions for constructive participation to appear. The corresponding rate-based sensitivity checks
are described with the statistical analyses.




% \subsubsection{User engagement}

% The unit of annotation is the conversational turn. User turns are labelled for behavioural, emotional and cognitive engagement, drawing on engagement theory and the ICAP distinction between passive, active and constructive participation~\cite{fredricks2004school,chi2014icap}. Passive engagement (\texttt{P}) captures reception or acknowledgement without substantial transformation. Active engagement (\texttt{A}) captures task-relevant action, follow-up or manipulation of supplied information. Constructive engagement (\texttt{C}) captures deeper work with ideas, such as testing an implication, revising a hypothesis, articulating a constraint, transferring a mechanism or extending an explanation. Visible emotional expression is labelled separately and reported descriptively because it was rare.

% The primary user-side outcome is constructive engagement, treated as the highest-level behavioural proxy for learning-oriented participation available in these logs. At the turn level, constructive engagement is a binary indicator for whether the user turn is labelled \texttt{C}. At the conversation level, we use both the count of constructive user turns and an indicator for whether a conversation contains at least one constructive user turn. Constructive ratios are defined as constructive user turns divided by user turns.


% \subsubsection{User engagement}

% The unit of annotation is the conversational turn. User turns are labelled with a depth hierarchy for uptake of assistant output, drawing on engagement theory and the ICAP distinction between passive, active and constructive participation
% ~\cite{fredricks2004school,chi2014icap}. Passive engagement (\texttt{P}) captures reception, acknowledgement or minimal continuation without substantial transformation. Active engagement (\texttt{A}) captures task-relevant use of, or
% follow-up on, supplied information, such as applying an answer, asking for a reformatted response or requesting the next step. Constructive engagement (\texttt{C}) captures articulated reasoning with generated output, such as testing an implication, revising a hypothesis or assumption, articulating a constraint, transferring a mechanism or extending an explanation. A user turn labelled \texttt{P}, \texttt{A} or \texttt{C} contributes to the overall cognitive-engagement
% measure.

% Explicit affective expression is labelled separately when a user turn expresses affect in text, and is not part of the passive--active--constructive depth hierarchy. We report
% this label descriptively to characterize the affective profile of the coding and writing exchanges. Operational definitions, plain-language examples and validation metrics
% are provided in Supplementary Tables~\ref{tab:annotation_codebook_definitions}--\ref{tab:production_label_validation}.

% Constructive engagement is the focal user-side measure in the main analyses because it identifies where generated output becomes material for the user's own reasoning,
% testing and revision. We represent it as a turn-level \texttt{C} indicator, a conversation-level count, an indicator for at least one constructive user turn and a constructive-turn ratio. The mapping from these measures to the descriptive,
% conversation-level, support-form and adjacent-turn analyses is summarized in Supplementary Tables~\ref{tab:model_specs}--\ref{tab:temporal_estimands}.

\subsubsection{User engagement}

The unit of annotation is the conversational turn. User turns are labelled with a depth hierarchy for uptake of assistant output, drawing on engagement theory and the ICAP distinction between passive, active and constructive participation
~\cite{fredricks2004school,chi2014icap}. Passive engagement (\texttt{P}) captures reception, acknowledgement or minimal continuation without substantial transformation. Active engagement (\texttt{A}) captures task-relevant use of, or
follow-up on, supplied information, such as applying an answer, asking for a reformatted response or requesting the next step. Constructive engagement (\texttt{C}) captures articulated reasoning with generated output, such as testing an
implication, revising a hypothesis or assumption, articulating a constraint, transferring a mechanism or extending an explanation. A user turn labelled \texttt{P}, \texttt{A} or \texttt{C} contributes to the overall cognitive-engagement
measure. Explicit affective expression is labelled in parallel when a user turn expresses affect in text; this label characterizes the affective profile of the exchange
alongside the cognitive-depth hierarchy. Operational definitions, plain-language examples and validation metrics are provided in Supplementary Tables~\ref{tab:annotation_codebook_definitions}--\ref{tab:production_label_validation}.

Constructive engagement is the focal user-side measure in the main analyses because it identifies where generated output becomes material for the user's own reasoning,
testing and revision. We represent it as a turn-level \texttt{C} indicator, a conversation-level count, an indicator for at least one constructive user turn and a
constructive-turn ratio. The mapping from these measures to the descriptive, conversation-level, support-form and adjacent-turn analyses is summarized in Supplementary Tables~\ref{tab:model_specs}--\ref{tab:temporal_estimands}.

% \subsubsection{Assistant support}

% Assistant turns are labelled for scaffolding. Scaffolded support denotes assistant behaviour that structures, regulates, diagnoses or redirects the user's work while leaving some responsibility with the user; the comparison category is a non-scaffolded reference response, including straightforward task fulfilment or stand-alone information provision. This definition follows the scaffolding tradition, especially contingency, transfer of responsibility and potential fading~\cite{wood1976role,van2010scaffolding,quintana2018scaffolding}.

% The primary assistant-side exposure is scaffolding, operationalized as scaffolded support. At the turn level, this is the distinction between scaffolded support and a non-scaffolded reference response. At the conversation level, we use an indicator for whether a conversation contains at least one scaffolded assistant turn. Scaffolded turns can also receive support-intent labels (\texttt{I1} metacognitive, \texttt{I2} cognitive and \texttt{I3} affective support) and non-mutually exclusive support-form labels (\texttt{M1} feedback, \texttt{M2} hinting, \texttt{M3} instructing, \texttt{M4} explaining, \texttt{M5} modelling and \texttt{M6} questioning). Support-form analyses condition on scaffolded turns and compare conversations containing each support-form label with scaffolded conversations that do not contain that label. Operational definitions are provided in Supplementary Table~\ref{tab:annotation_codebook_definitions}; paraphrased examples and two complete de-identified case transcripts are provided in Supplementary Tables~\ref{tab:label_examples}--\ref{tab:complete_writing_case}.

\subsubsection{Assistant support}

Assistant turns are labelled by their role in the user's ongoing work. The top-level support distinction separates reference responses from scaffolded support. Reference
responses primarily provide the requested information, deliverable or stand-alone answer. Scaffolded support captures assistant behaviour that structures, regulates, diagnoses or redirects the user's process while preserving a role for the user's own reasoning. This definition follows the scaffolding tradition, especially contingency, transfer of responsibility and potential fading~\cite{wood1976role,van2010scaffolding,quintana2018scaffolding,reiser2004scaffolding}.

Scaffolded support is the focal assistant-side measure in the main analyses. At the turn level, assistant turns are represented as scaffolded support or reference response.
At the conversation level, we use an indicator for whether a conversation contains at least one scaffolded assistant turn. Scaffolded turns are further characterized by support-intent labels (\texttt{I1} metacognitive, \texttt{I2} cognitive and
\texttt{I3} affective support) and non-mutually exclusive support-form labels (\texttt{M1} feedback, \texttt{M2} hinting, \texttt{M3} instructing, \texttt{M4} explaining, \texttt{M5} modelling and \texttt{M6} questioning). The intent labels describe the broad purpose of support, whereas the form labels
describe how support is delivered in the turn. Support-form analyses condition on scaffolded turns and compare scaffolded
conversations containing each support-form label with scaffolded conversations that do not contain that label. Because support-form labels are multi-label descriptors,
these contrasts trace how particular forms of scaffolding pattern with constructive participation within the broader scaffolded-support ecology.

Operational definitions and plain-language examples are provided in Supplementary Tables~\ref{tab:annotation_codebook_definitions} and \ref{tab:label_examples}; the complete de-identified case transcripts are provided in Supplementary Tables~\ref{tab:complete_coding_case}--\ref{tab:complete_writing_case}. Validation metrics for scaffolded support and support-form labels are reported in Supplementary Tables~\ref{tab:label_validation} and \ref{tab:production_label_validation}. The mapping
from assistant-side measures to conversation-level, support-form and adjacent-turn analyses is summarized in Supplementary
Tables~\ref{tab:model_specs}--\ref{tab:temporal_estimands}.


% \subsection{Annotation workflow and validation}

% The annotation workflow combined theory-grounded codebook development, prompt iteration, parser checks, targeted post-processing, human review and production-scale LLM-assisted labelling. We used LLM-assisted annotation as a scalable text-analysis workflow while retaining human verification because prior work shows both the promise and the need for validation when language models are used as annotators~\cite{gilardi2023chatgpt,ziems2024can}. After the workflow stabilized, the same construct definitions, schema checks and post-processing logic were applied to the WildChat, LMSYS Chat and ShareChat analytic corpora. The archived annotation metadata record the provider, deployment family, API version, prompt version, retry policy, parser checks and rule-based post-processing rules used for production labelling; Supplementary Section A.3 gives the reproducibility details.

% Human verification is reported in layers. Human--human agreement assesses whether trained reviewers could apply the codebook consistently. Human--LLM production-label checks compare final or post-processed production labels with human-confirmed labels. Targeted post-production audits check whether narrow rule-based reviews reduced known residual false positives. Agreement is summarized with per-label F1, Matthews correlation coefficient and Gwet's AC1 because the labels are sparse, prevalence-imbalanced and sometimes non-mutually exclusive~\cite{matthews1975comparison,gwet2008computing}. The consolidated production-label audits indicate acceptable or better reliability across the labels used for the main claims: intentional framing reached F1=0.857, constructive-versus-non-constructive user engagement reached F1=0.922, broad scaffolded support reached F1=0.912 and support-form labels ranged from F1=0.691 to 0.938. Supplementary Tables~\ref{tab:label_validation} and \ref{tab:production_label_validation} report the complete validation metrics and audit designs.

\iffalse
Archived previous Methods 4.4 version before redundancy reduction.

\subsection{Annotation workflow and validation}

The source corpora do not contain native labels for learning-oriented engagement, assistant scaffolding, support intent or support form. We therefore constructed these labels through an iterative annotation workflow that combined theory-grounded
codebook development, independent human review, prompt calibration, parser checks, targeted post-processing and production-scale LLM-assisted labelling.

\paragraph{Codebook development and domain adaptation.}
The codebook was drafted from learning-science and scaffolding constructs: user-side engagement followed engagement theory and the ICAP distinction between passive, active and constructive participation~\cite{fredricks2004school,chi2014icap}, whereas
assistant-side support followed scaffolding accounts emphasizing contingency, structuring of activity and transfer of responsibility~\cite{wood1976role,van2010scaffolding,quintana2018scaffolding,reiser2004scaffolding}.
The initial operational codebook was developed in the coding task ecology, where errors, execution failures, repair attempts and iterative testing provided dense boundary
cases for distinguishing user uptake depth and assistant support. Two trained author-reviewers independently labelled overlapping coding validation cases across three rounds, reviewed disagreements and produced consensus or adjudicated
human-confirmed labels. These coding calibration rounds covered 658 task turns, and refined recurring boundaries including passive versus active uptake, active versus constructive engagement, direct answer delivery versus scaffolded support, explicit affective expression and co-occurring support-form labels.

After the coding codebook stabilized, we adapted the same construct categories to writing. The adaptation preserved the passive--active--constructive user-engagement hierarchy, the reference-response versus scaffolded-support distinction, and the support-intent and support-form label sets, but added writing-specific decision rules and few-shot examples for drafting, rewriting, tone adjustment, evaluative choice and rationale articulation. We then conducted a writing-specific validation pass on 210 reviewed conversations to verify that the adapted codebook could be applied to
writing exchanges. These review rounds anchored the final operational boundaries before production-scale annotation. The final operational definitions are reported in
Supplementary Table~\ref{tab:annotation_codebook_definitions}; paraphrased label examples and two complete de-identified case transcripts are provided in Supplementary Tables~\ref{tab:label_examples}- \ref{tab:complete_writing_case}. Human--human agreement metrics for this codebook-operability layer are reported in Supplementary Table~\ref{tab:label_validation}, and Supplementary Section~A.4 describes the review-round sizes and adjudication workflow.

\paragraph{Prompt calibration and production labelling.}
LLM-assisted annotation was used as the scalable production procedure after the domain-adapted codebook, prompt templates and structured output schemas had been calibrated against human-reviewed examples. Prompt outputs were first tested on a
label-enriched pilot calibration set of 400 conversations, stratified by task ecology with 200 coding and 200 writing conversations. The pilot was assembled from preliminary LLM labels across all annotation prompt families, with quotas requiring at least 30 candidate turn-level instances for every target label, and was supplemented with randomly sampled background conversations without target positive labels.
Because sampling was conducted at the conversation level but label quotas were defined at the turn level, the 400 conversations yielded a substantially larger set of reviewed
turns for testing label boundaries.

Prompt templates and schemas were revised against observed disagreement patterns, schema failures and boundary errors. After these revisions, a balanced 200-turn human validation check, with 100 coding and 100 writing examples, was used to test the refreshed user-turn label boundaries and identify remaining recurring false-positive patterns. The resulting updates were incorporated into the frozen codebook--schema--parser--post-processing workflow used for final analysis labels.
After the WildChat production labelling pass, we also reviewed 100-turn post-labelling samples from each WildChat task setting, with 50 user turns and 50 assistant turns per task drawn from the completed labelled pools under the observed production-label distribution. WildChat labels reflect the targeted update pass applied after validation; LMSYS Chat and ShareChat were then processed with the same refreshed workflow. Production labelling used the same construct definitions, prompt versions, schema checks and
post-processing logic across corpora.

\paragraph{Validation design.}
Validation evidence was mapped to the stage of the workflow it evaluated. The first layer assessed codebook operability through human--human agreement. This layer tested whether trained reviewers could apply the finalized construct definitions after coding codebook development and writing-domain adaptation. The coding codebook-operability check combined three validation rounds covering 658 task turns. The writing-domain validation used 210 reviewed conversations after the coding-derived codebook had been adapted with writing-specific decision rules and few-shot examples. Human--human agreement for this layer is reported in
Supplementary Table~\ref{tab:label_validation}; Supplementary Section~A.4 describes the review rounds, adjudication process and domain-adaptation workflow.

The second layer assessed production-label quality after the frozen codebook--schema--parser--post-processing workflow had been applied. These production-label audits compared final LLM-assisted labels with human-confirmed labels and were stratified across the six corpus-by-task settings. The user-framing
audit reviewed 450 first user turns, with 75 cases per setting: 50 production-positive and 25 production-negative cases. The user-engagement audit reviewed 600 user turns, with 100 cases per setting and quota enrichment by production label: 50 constructive, 20 active, 15 passive and 15 no-engagement cases. The assistant-side audit reviewed 180 assistant turns, with 30 cases per setting: 24 production-scaffolded and 6 reference-response cases. Support-form positives were enriched within the
scaffolded cases, and support-form labels were treated as non-exclusive. Production-label audit metrics are reported in Supplementary Table~\ref{tab:production_label_validation}.

The third layer consisted of targeted post-production checks for recurring boundary cases identified during validation and prompt calibration. A final user-turn check reviewed 100 coding and 100 writing examples after the prompt and boundary-rule
updates. After the whole process, the WildChat analytic sample contained 31,878 coding conversations and 39,534 writing conversations. LMSYS Chat and ShareChat were processed with the same refreshed workflow. Supplementary Section~A.4 reports the post-update checks and sampling details, and Supplementary Table~\ref{tab:filter_classifier_validation} summarizes validation and audit evidence for task ecology, language filtering, user framing and turn-level labels.

\paragraph{Agreement metrics.}
Agreement is summarized with per-label F1, Matthews correlation coefficient and Gwet's AC1 because the labels are sparse, prevalence-imbalanced and sometimes multi-label~\cite{matthews1975comparison,gwet2008computing}. We report agreement separately for the two main validation layers. Supplementary
Table~\ref{tab:label_validation} reports human--human codebook agreement, which evaluates whether trained reviewers could apply the operational definitions consistently
during codebook calibration and writing-domain adaptation. Supplementary Table~\ref{tab:production_label_validation} reports production-label agreement, which evaluates the final LLM-assisted labels after schema checks, parser checks and
post-processing.

The consolidated production-label audits support the labels used for the main claims. Intentional user framing reached F1=0.857; constructive-versus-non-constructive user
engagement reached F1=0.922; broad scaffolded support reached F1=0.912; and support-form labels ranged from F1=0.691 to 0.938. The audit samples intentionally enriched positive and boundary-relevant cases to stress-test classification boundaries; prevalence estimates are computed from the full production labels reported in the Results~\ref{sec:results}.
\fi

\subsection{Annotation workflow and validation}

The source corpora do not contain native labels for learning-oriented engagement, assistant scaffolding, support intent or support form. We therefore constructed these labels through an iterative workflow that combined theory-grounded codebook development, independent human review, prompt calibration, parser checks, targeted post-processing and production-scale LLM-assisted labelling~\cite{gilardi2023chatgpt,ziems2024can}. The codebook was derived from engagement theory, the ICAP framework and scaffolding theory~\cite{fredricks2004school,chi2014icap,wood1976role,van2010scaffolding,quintana2018scaffolding,reiser2004scaffolding}. It was developed first in coding-oriented conversations, where errors and repair attempts provided dense boundary cases, and then adapted to writing-oriented conversations with writing-specific decision rules and examples. Human reviewers independently labelled overlapping validation cases, discussed disagreements and produced human-confirmed labels for validation analyses.

After the codebook, prompt templates and structured output schemas stabilized, the frozen annotation workflow was applied to the analytic corpora. The workflow used schema and parser checks, documented post-processing rules for recurring boundary errors and retained annotation metadata for provider, deployment family, API version, prompt version, retry policy and parser checks. WildChat labels reflect the targeted update pass applied after validation, including post-labelling WildChat review samples drawn from completed labelled pools; LMSYS Chat and ShareChat were then processed with the same refreshed workflow. Operational definitions, examples, complete de-identified cases and reproducibility details are provided in Supplementary Tables~\ref{tab:annotation_codebook_definitions}--\ref{tab:complete_writing_case} and Supplementary Sections A.3--A.4.

Validation evidence is reported in three layers. Human--human agreement assesses codebook operability; production-label audits compare final LLM-assisted labels with human-confirmed labels; and post-update audits check recurring residual boundary errors. Agreement is summarized with per-label F1, Matthews correlation coefficient and Gwet's AC1 because the labels are sparse, prevalence-imbalanced and sometimes multi-label~\cite{matthews1975comparison,gwet2008computing}. The consolidated production-label audits support the labels used for the main claims: intentional user framing reached F1=0.857, constructive-versus-non-constructive user engagement reached F1=0.922, broad scaffolded support reached F1=0.912 and support-form labels ranged from F1=0.691 to 0.938. Full sampling designs, review-round sizes, post-labelling checks and per-label metrics are reported in Supplementary Section A.4 and Supplementary Tables~\ref{tab:label_validation}--\ref{tab:wildchat_post_labelling_review}. Prevalence estimates in the Results are computed from the full production labels.


% \subsection{Statistical analyses}

% Descriptive analyses estimate engagement prevalence, scaffolding prevalence, passive/active/constructive composition, visible emotional-expression rates and behavioural-depth summaries within each task setting. Context analyses compare observed user framing, task ecology and depth groups. Because the probability that a conversation contains at least one constructive turn partly reflects the number of turns available for observing one, we also fit a grouped-binomial constructive-rate sensitivity model in which the outcome is the constructive user-turn count out of total user turns, with user framing, task ecology, length bucket and dataset fixed effects as predictors.

% Conversation-level support associations compare conversations with and without at least one scaffolded assistant turn. Constructive-ratio contrasts use turn-weighted ratios within each comparison group, with uncertainty from 1,000 conversation-level bootstrap resamples. Adjusted Section~2.3 models use setting-specific Poisson generalized linear models for raw constructive-turn counts and companion logistic models for whether a conversation contains at least one constructive user turn. The primary Poisson model treats conversation length as a covariate rather than using a user-turn offset; offset-rate and quasi-Poisson sensitivities are reported in the Supplement. These models are interpreted as adjusted associations, not causal effects.

% Support-form contrasts compare scaffolded conversations containing a given support form with scaffolded conversations not containing that form. Because support-form labels are non-mutually exclusive and nested within scaffolded support, these contrasts are descriptive form-pattern analyses rather than mutually exclusive treatment comparisons. Bracket tests in Fig.~\ref{fig:support_form_supply} use Benjamini--Hochberg false-discovery-rate adjustment within each displayed family of six support-form comparisons.

% Temporal analyses focus on adjacent-turn ordering. We compare the probability that the next user turn is constructive after scaffolded versus non-scaffolded assistant turns, estimate the same contrast within prior user states and estimate the probability that the next assistant turn is scaffolded after different prior user states. Adjacent-turn contrasts use 1,000 conversation-cluster bootstrap resamples because multiple turn pairs can come from the same conversation. Integrated adjacent-turn regressions combine user-state features, assistant-scaffolding features and adjacent-turn context in one model, with conversation-clustered standard errors. Supplementary Tables~\ref{tab:model_specs}--\ref{tab:temporal_estimands} and \ref{tab:setting_model_covariates} give the complete estimand map, covariate sets, model specifications and sensitivity checks.

\subsection{Statistical analyses}

\paragraph{Descriptive and contextual estimates.}
Descriptive analyses estimated engagement prevalence, scaffolding prevalence, passive/active/constructive composition, explicit affective-expression rates and
interaction-depth summaries within each task setting. Contextual analyses compared observed user framing, task ecology and interaction-depth groups. For analyses using
whether a conversation contained at least one constructive user turn, we paired the event indicator with a rate-based sensitivity model to account for the number of user turns available for observing constructive participation. The sensitivity model was a grouped-binomial model in which the outcome was the constructive user-turn count out of total user turns, with user framing, task ecology, length bucket and dataset fixed effects as predictors.

\paragraph{Conversation-level support associations.}
Conversation-level support analyses compared conversations containing at least one scaffolded assistant turn with reference conversations without scaffolded support.
Constructive-ratio contrasts used turn-weighted ratios within each comparison group, defined as total constructive user turns divided by total user turns, with uncertainty estimated from 1,000 conversation-level bootstrap resamples. Covariate-adjusted
models reported in Section~2.3 used setting-specific Poisson generalized linear models for raw constructive-turn counts and companion logistic models for whether a conversation contained at least one constructive user turn~\cite{mccullagh1989generalized,agresti2013categorical}.
The primary Poisson model included conversation length as a covariate; offset-rate and quasi-Poisson sensitivities are reported in the Supplement. The resulting estimates are interpreted as covariate-adjusted associations in naturalistic conversation logs.

\paragraph{Support-form analyses.}
Support-form analyses conditioned on scaffolded conversations. For each support-form label, we compared scaffolded conversations containing that label with scaffolded conversations that did not contain that label, and reported turn-weighted constructive user-turn ratio differences. Because support-form labels are multi-label descriptors, these contrasts trace how particular forms of scaffolding pattern with
constructive participation within the broader scaffolded-support ecology. Bracket tests in Fig.~\ref{fig:support_form_supply} compare displayed strata for each support form; q values use Benjamini--Hochberg false-discovery-rate adjustment within each displayed family of six support-form comparisons~\cite{benjamini1995controlling}.

\paragraph{Adjacent-turn analyses.}
Temporal analyses focused on adjacent-turn ordering. We compared the probability that the next user turn was constructive after scaffolded versus reference assistant
turns, estimated the same contrast within prior user states and estimated the probability that the next assistant turn was scaffolded after different prior user states. Adjacent-turn contrasts used 1,000 conversation-cluster bootstrap resamples because multiple turn pairs can come from the same conversation.

Integrated adjacent-turn regressions combined prior user-state features, assistant-scaffolding features, support-form descriptors and adjacent-turn context in one logistic model predicting whether the next user turn was constructive. These
models included task and framing variables, turn position and dataset fixed effects; a sensitivity replaced dataset fixed effects with recoverable model/source-family fixed effects where available. Standard errors were clustered by conversation. We report broad scaffolded-support models separately from support-form decompositions because
support-form labels are descriptors nested within scaffolded support. Nested likelihood-ratio block tests first add broad scaffolded-support presence and then test whether support-form descriptors add signal within scaffolded support. Supplementary
Tables~\ref{tab:model_specs}--\ref{tab:temporal_estimands} and
\ref{tab:setting_model_covariates} give the complete estimand map, covariate sets, model specifications and sensitivity checks.

% \subsection{LLM use and reporting boundaries}

% LLMs were used as objects of study and as annotation tools. The authors also used LLM-based tools to assist with code drafting, debugging, analysis workflow organization and language editing. LLM-generated assistance was not used as a substitute for human verification of the validation samples or for interpretation of the statistical results.

% All reported results are aggregate turn-level or conversation-level analyses. The study does not link conversations into user histories, estimate user-level longitudinal trajectories or redistribute verbatim raw conversation logs. Operational examples are de-identified and paraphrased or lightly edited, and temporal analyses are confined to ordering within a conversation. Model/source checks are used as robustness analyses rather than model-causal comparisons, because model labels in public logs are not experimentally assigned and may be confounded with time, user population, task mix and logging context.




\subsection{LLM use and reporting boundaries}

Large language models were used in four roles. First, LLMs were the systems under investigation: the data are public human--LLM conversations from WildChat, LMSYS Chat and ShareChat. Available model/source identifiers were retained as observational metadata for descriptive and robustness analyses, not as randomized model comparisons.

Second, LLMs were used as measurement tools. GPT-4o-family Azure OpenAI deployments were used for coding-versus-writing task filtering and English-oriented filtering. Azure OpenAI GPT-5.1-family deployments were used for production annotation:
\texttt{gpt-5.1} for the main user-turn and assistant-turn labels, and
\texttt{gpt-5.1-chat} for final writing-oriented annotation. These calls assigned user framing, engagement, scaffolding, support intent and support-form labels under fixed prompts, schema/parser checks, retry logic, post-processing rules and human validation.

Third, LLMs were used during annotation development to test prompts and inspect boundary cases against human-reviewed examples; pilot outputs were not used as production labels or prevalence estimates. Fourth, LLM-based tools assisted with code drafting, debugging, workflow organization and language editing, but not as substitutes for human validation, audit-label adjudication, statistical decisions or result interpretation.

All analyses are aggregate turn- or conversation-level analyses. Model/source labels are observational descriptors, so model-related comparisons are interpreted as robustness analyses rather than causal effects.
