\section*{Data availability}

This study analyses public human--LLM conversation corpora released by their original data providers: WildChat, LMSYS Chat-1M and ShareChat. Readers should obtain the raw public conversation data from the original releases and comply with the corresponding dataset licences, terms of use and redistribution restrictions. The manuscript does not redistribute verbatim raw conversation logs, user identifiers or linked user histories. The Supplementary includes two de-identified illustrative case excerpts derived from retained conversations and formatted to preserve turn sequence, task structure and annotation decisions while reducing privacy and licence risk.

Derived data sufficient to verify the reported aggregate figures, tables and statistical summaries are archived on Zenodo at \url{https://doi.org/10.5281/zenodo.20995945}. The same release is also available in the GitHub repository at \url{https://github.com/CinderD/Informal-learning-in-everyday-human-LLM-interaction}. The release includes the numeric values underlying the main figures and Appendix C visual summaries, de-identified conversation-level and adjacent-turn analytic labels without raw message text, table source files, confidence intervals, p values, false-discovery-rate q values, model-output summaries, provenance documentation and SHA-256 checksums.

\section*{Code availability}

Custom code used to export source data and de-identified label tables, compute statistical analyses and generate figures and tables is archived with the derived-data release on Zenodo at \url{https://doi.org/10.5281/zenodo.20995945} and is also available in the GitHub repository at \url{https://github.com/CinderD/Informal-learning-in-everyday-human-LLM-interaction}. \revise{The archived repository includes source-data export scripts, statistical-analysis scripts, figure-generation scripts, table source files and annotation reproducibility summaries covering prompt families, output fields, parser checks, retry logic and rule-based post-processing. API keys, raw message text and non-redistributable corpus files are excluded, so full raw-to-label reruns require user-provided credentials and access to the original public corpora.}

\section*{Ethics statement}

The study is an analysis of public, naturalistic human--LLM conversation datasets. 
We used the public source corpora as released by their original providers, with provider-side anonymization or de-identification already applied. Reporting is therefore restricted to aggregate statistics, source-data summaries, validation metrics and two de-identified illustrative case excerpts that preserve turn order and annotation decisions.
Human validation was used to assess annotation quality on reviewed examples and did not involve interaction with the original dataset users.

% The authors did not recruit participants, intervene in conversations or attempt to identify users. Analyses are reported at aggregate turn or conversation level. \revise{Operational examples are de-identified and paraphrased; the two illustrative case excerpts preserve turn structure and annotation decisions but are not redistributed verbatim raw logs.} Human validation was used to assess annotation quality on reviewed examples and did not involve interaction with the original dataset users.

\section*{Competing interests}

H.L. and X.X. are employees of Microsoft Research Asia. The remaining authors declare no competing interests.

\section*{Author contributions}

% Author contributions are organized using the CRediT taxonomy. The contribution categories for this study are conceptualization, study design, data curation, annotation pipeline development, validation, statistical analysis, figure generation, writing of the original draft, manuscript review and editing, supervision and funding acquisition. Author initials and final role allocation: [to be added after the author list and order are confirmed]. All listed authors will approve the submitted version and will be accountable for their contributions.

Z.C. and H.L. conceived the project. H.L., X.X. and H.Q. secured funding for the project. Z.C. and H.L. collected the data. Z.C. and H.L. managed the data processing and analysis process. Z.C. and H.L. prepared the analysis of the results. Z.C. drafted the manuscript. All authors contributed to the interpretation of the results and edited and revised the manuscript. H.L. and X.X. supervised the project.

% T.H., J.H.-O., D.S. and L.S. conceived the project. X.X., D.S. and L.S. secured funding for the project. T.H., Y.J. and L.S. collected the data. H.L. and L.S. managed
% the creativity evaluation process. T.H. prepared the result analysis and plotting and
% drafted the manuscript. All authors contributed to the interpretation of the results
% and edited and revised the manuscript. D.S. and L.S. supervised the project.

\section*{Corresponding authors}
Correspondence to Haotian Li.


\section*{Acknowledgements and funding}

TBD.

% Acknowledgements and funding information, including grant numbers where applicable: [to be added]. Funders had no role in study design, data collection, analysis, interpretation, manuscript preparation or publication decisions unless stated otherwise.

\section*{LLM use statement}

\todayrevise{Large language models were used in four roles. First, LLMs were the systems under study: the analysed data consist of public human--LLM conversations from WildChat, LMSYS Chat and ShareChat. Available assistant-model or service-source identifiers were retained as observational metadata for descriptive and robustness analyses. Second, LLMs were used as measurement tools for task filtering, English-oriented filtering and corpus-scale annotation of user framing, engagement, scaffolding, support intent and support form. These annotation workflows used fixed prompts, schema and parser checks, retry logic, post-processing rules and human validation; the reported reproducibility materials summarize model or deployment families and implementation controls in Supplementary Section~\hyperref[supp:a3]{A.3}. Third, LLMs were used during annotation development to test prompts and inspect boundary cases against human-reviewed examples. Fourth, LLM-based tools assisted with code drafting, debugging, workflow organization and language editing. LLM-generated assistance was not used as a substitute for human validation, audit-label adjudication, statistical decisions or interpretation of the results. No LLM or AI tool is listed as an author, and the authors take responsibility for the final manuscript, analyses and interpretation.}

\section*{Figure source data}
{\color{blue}
\begin{tabularx}{\textwidth}{@{}>{\raggedright\arraybackslash}p{0.20\textwidth}>{\raggedright\arraybackslash}X@{}}
Main figures & Fig.~2: \path{source_data/figure2_source_data.csv}; Fig.~3: \path{source_data/figure3_source_data.csv}; Fig.~4: \path{source_data/figure4_source_data.csv}; Fig.~5: \path{source_data/figure5_source_data.csv}. \\
Fig.~1 & Conceptual framework figure; no numeric source-data file. \\
Appendix displays and tables & Source files for Appendix C visual summaries, supplementary tables and regression outputs are provided under \path{source_data/}, \path{tables/} and \path{statistical_outputs/}. \\
\end{tabularx}
}

% Source data for the numeric main figures are provided in \texttt{source\_data/figure2\_source\_data.csv}, \texttt{source\_data/figure3\_source\_data.csv}, \texttt{source\_data/figure4\_source\_data.csv} and \texttt{source\_data/figure5\_source\_data.csv}. Figure~1 is a conceptual framework figure and has no numeric source data. Supplementary table and regression source files are provided in \texttt{tables/} and \texttt{outputs/integrated\_regression/}.
