\begin{sidewaystable}[p]
\fontsize{7.0pt}{8.2pt}\selectfont
\centering
\caption{\textbf{Corpus filtering ledger.} Counts show the corpus-specific steps from public source records to the final coding- and writing-oriented analytic samples. Filter validation evidence is reported separately in Supplementary Table~\ref{tab:filter_classifier_validation}.}
\label{tab:corpus_filtering_pipeline}
\setlength{\tabcolsep}{4pt}
\renewcommand{\arraystretch}{1.12}
\begin{tabularx}{\textheight}{>{\raggedright\arraybackslash}p{0.12\textheight}>{\raggedright\arraybackslash}p{0.20\textheight}>{\raggedright\arraybackslash}p{0.19\textheight}>{\raggedright\arraybackslash}p{0.30\textheight}>{\raggedright\arraybackslash}X}
\toprule
Corpus
& Public source records
& Eligibility screen
& Task and language filtering
& Final analytic sample \\
\midrule
WildChat
& WildChat-4.8M public release.
& English-oriented records with at least four message turns.
& Keyword prefilter followed by LLM-assisted semantic filters: \texttt{coding\_english} or \texttt{writing\_english}, score $\geq .70$. One coding record with missing \texttt{learning\_intent} was excluded.
& 31,878 coding; 39,534 writing. \\
\addlinespace
LMSYS Chat
& LMSYS Chat-1M public release: 1,000,000 conversations.
& 777,453 English conversations; 236,372 English conversations with at least four turns.
& LLM-assisted semantic filters applied to the English pool with the minimum-turn screen: 32,114 coding; 21,023 writing.
& 32,114 coding; 21,023 writing. \\
\addlinespace
ShareChat
& ShareChat public release: 142,808 public conversations and 660,293 public turns. Local conversion used 1,199,899 message rows and 129,584 platform-url conversations.
& 69,915 conversations with at least four message turns.
& LLM-assisted semantic filters selected 3,494 coding and 1,663 writing conversations; the final strict-English screen retained 2,481 coding and 1,539 writing.
& 2,481 coding; 1,539 writing. \\
\bottomrule
\end{tabularx}

\vspace{2mm}
\begin{minipage}{0.95\textheight}
\footnotesize
\textit{Note.} Counts use corpus-specific source units. The minimum-turn screen uses message count, not user--assistant round count. For ShareChat, public turns and local message rows are different units. ``Strict-English'' refers only to the additional ShareChat final-language screen.
\end{minipage}
\end{sidewaystable}
